---
title: "General Decision Tree Model"
output: pdf_document
---

## Overview

This document builds a decision tree classifier that is tuned to favor generalization rather than memorization.  
A moderate cross-validation grid is used to balance the model tuning and run time

```{r setup, message=FALSE, warning=FALSE}
packages <- c("data.table", "rpart", "rpart.plot", "caret", "ggplot2")

installed <- rownames(installed.packages())
for (p in packages) {
  if (!(p %in% installed)) install.packages(p)
}

library(data.table)
library(rpart)
library(rpart.plot)
library(caret)
library(ggplot2)

set.seed(571)
```

## Load the data

Make sure the CSV files are inside a folder called `data` in your project directory.

```{r load-data}
train <- fread("data/dota2Train.csv", header = FALSE)
test  <- fread("data/dota2Test.csv", header = FALSE)

colnames(train)[1:4] <- c("result", "clusterid", "gamemode", "gametype")
colnames(test)[1:4]  <- c("result", "clusterid", "gamemode", "gametype")

hero_names <- paste0("hero", 1:(ncol(train) - 4))
colnames(train)[5:ncol(train)] <- hero_names
colnames(test)[5:ncol(test)]   <- hero_names

train$result <- factor(train$result, levels = c(-1, 1), labels = c("loss", "win"))
test$result  <- factor(test$result, levels = c(-1, 1), labels = c("loss", "win"))
```

## Cross-validation over tree settings

To reduce runtime, we search a smaller grid:
- `maxdepth` values from 3 to 12
- `cp` values of 0.0005, 0.001, and 0.005
- 3-fold cross-validation

```{r cv-search}
depth_grid <- 3:12
cp_grid <- c(0.0005, 0.001, 0.005)

folds <- createFolds(train$result, k = 3, returnTrain = FALSE)

results <- data.frame()

for (d in depth_grid) {
  for (cp_val in cp_grid) {

    fold_acc <- c()

    for (i in seq_along(folds)) {
      val_idx <- folds[[i]]
      train_fold <- train[-val_idx]
      val_fold   <- train[val_idx]

      tree_fit <- rpart(
        result ~ .,
        data = train_fold,
        method = "class",
        control = rpart.control(
          maxdepth = d,
          minsplit = 20,
          minbucket = 10,
          cp = cp_val
        )
      )

      pred_val <- predict(tree_fit, val_fold, type = "class")
      acc <- mean(pred_val == val_fold$result)
      fold_acc <- c(fold_acc, acc)
    }

    results <- rbind(
      results,
      data.frame(
        maxdepth = d,
        cp = cp_val,
        cv_accuracy = mean(fold_acc)
      )
    )
  }
}

best_row <- results[which.max(results$cv_accuracy), ]
best_row
```

## Plot CV accuracy across all trees

```{r cv-plot, fig.width=10, fig.height=6}
ggplot(results, aes(x = maxdepth, y = cv_accuracy, color = as.factor(cp), group = cp)) +
  geom_line() +
  geom_point() +
  geom_point(
    data = best_row,
    aes(x = maxdepth, y = cv_accuracy),
    inherit.aes = FALSE,
    size = 4
  ) +
  labs(
    title = "Cross-Validated Accuracy Across Tree Settings",
    x = "Max Depth",
    y = "CV Accuracy",
    color = "cp"
  )
```

## Fit the best general tree

```{r fit-best-general}
best_tree <- rpart(
  result ~ .,
  data = train,
  method = "class",
  control = rpart.control(
    maxdepth = best_row$maxdepth,
    minsplit = 20,
    minbucket = 10,
    cp = best_row$cp
  )
)

printcp(best_tree)
```

## Plot the best general tree

```{r plot-best-tree, fig.width=12, fig.height=8}
rpart.plot(best_tree, main = "Best General Decision Tree")
```

## Evaluate on training and test sets

```{r evaluate-general}
train_pred <- predict(best_tree, train, type = "class")
test_pred  <- predict(best_tree, test, type = "class")

train_cm <- confusionMatrix(data = train_pred, reference = train$result, positive = "win")
test_cm  <- confusionMatrix(data = test_pred, reference = test$result, positive = "win")

cat("Best hyperparameters:\n")
print(best_row)

cat("\nTraining accuracy:\n")
print(train_cm$overall["Accuracy"])

cat("\nTest accuracy:\n")
print(test_cm$overall["Accuracy"])

cat("\nTraining confusion matrix:\n")
print(train_cm$table)

cat("\nTest confusion matrix:\n")
print(test_cm$table)

cat("\nFull test confusion matrix statistics:\n")
print(test_cm)
```

## Notes

This model is intended to be the more generalizable tree:
- moderate depth
- moderate split restrictions
- limited pruning search for faster runtime
- evaluated on unseen test data
