---
title: "Overfitting Decision Tree Model"
output: pdf_document
---

## Overview

This document builds an intentionally overfit decision tree classifier.  
The goal is to allow the tree to grow with almost no pruning so that we can compare its behavior against the general model.

```{r setup, message=FALSE, warning=FALSE}
packages <- c("data.table", "rpart", "rpart.plot", "caret")

installed <- rownames(installed.packages())
for (p in packages) {
  if (!(p %in% installed)) install.packages(p)
}

library(data.table)
library(rpart)
library(rpart.plot)
library(caret)

set.seed(571)
```

## Load the data

Make sure the CSV files are inside a folder called `data` in your project directory.

```{r load-data}
train <- fread("data/dota2Train.csv", header = FALSE)
test  <- fread("data/dota2Test.csv", header = FALSE)

colnames(train)[1:4] <- c("result", "clusterid", "gamemode", "gametype")
colnames(test)[1:4]  <- c("result", "clusterid", "gamemode", "gametype")

hero_names <- paste0("hero", 1:(ncol(train) - 4))
colnames(train)[5:ncol(train)] <- hero_names
colnames(test)[5:ncol(test)]   <- hero_names

train$result <- factor(train$result, levels = c(-1, 1), labels = c("loss", "win"))
test$result  <- factor(test$result, levels = c(-1, 1), labels = c("loss", "win"))
```

## Fit the overfitting model

This tree is intentionally allowed to grow aggressively:
- very deep tree
- tiny split restrictions
- no practical pruning

```{r fit-overfit}
overfit_tree <- rpart(
  result ~ .,
  data = train,
  method = "class",
  control = rpart.control(
    maxdepth = 30,
    minsplit = 2,
    minbucket = 1,
    cp = 0
  )
)

printcp(overfit_tree)
```

## Plot a simplified display version

The full overfit tree is usually too dense to read clearly.  
So this section keeps the full overfit model for evaluation, but shows a lightly pruned copy for visualization only.

```{r plot-overfit-tree, fig.width=12, fig.height=8}
display_tree <- prune(overfit_tree, cp = 0.001)

rpart.plot(
  display_tree,
  main = "Simplified View of the Overfitting Tree",
  cex = 0.6,
  compress = TRUE,
  fallen.leaves = TRUE,
  faclen = 0,
  varlen = 0
)
```

## Evaluate on training and test sets

```{r evaluate-overfit}
train_pred <- predict(overfit_tree, train, type = "class")
test_pred  <- predict(overfit_tree, test, type = "class")

train_cm <- confusionMatrix(data = train_pred, reference = train$result, positive = "win")
test_cm  <- confusionMatrix(data = test_pred, reference = test$result, positive = "win")

cat("Training accuracy:\n")
print(train_cm$overall["Accuracy"])

cat("\nTest accuracy:\n")
print(test_cm$overall["Accuracy"])

cat("\nTraining confusion matrix:\n")
print(train_cm$table)

cat("\nTest confusion matrix:\n")
print(test_cm$table)

cat("\nFull test confusion matrix statistics:\n")
print(test_cm)
```

## Notes

This model is intended to overfit:
- extremely flexible structure
- almost no regularization
- expected to fit the training data much more closely
- may not improve test-set performance

The tree shown in the figure is a simplified display version so it fits in the PDF more clearly.
