---
title: "Overfitting Decision Tree Model"
output: pdf_document
---

## Overview

This document builds an intentionally overfit decision tree classifier.  
The full overfit tree is evaluated using confusion matrices and summary metrics.  
No rendered tree figure is included in this version.

```{r setup, message=FALSE, warning=FALSE}
packages <- c("data.table", "rpart", "caret")

installed <- rownames(installed.packages())
for (p in packages) {
  if (!(p %in% installed)) install.packages(p)
}

library(data.table)
library(rpart)
library(caret)

set.seed(571)
```

## Load the data

Make sure the CSV files are inside a folder called `data` in your project directory.

```{r load-data}
train <- fread("data/dota2Train.csv", header = FALSE)
test  <- fread("data/dota2Test.csv", header = FALSE)

colnames(train)[1:4] <- c("result", "clusterid", "gamemode", "gametype")
colnames(test)[1:4]  <- c("result", "clusterid", "gamemode", "gametype")

hero_names <- paste0("hero", 1:(ncol(train) - 4))
colnames(train)[5:ncol(train)] <- hero_names
colnames(test)[5:ncol(test)]   <- hero_names

train$result <- factor(train$result, levels = c(-1, 1), labels = c("loss", "win"))
test$result  <- factor(test$result, levels = c(-1, 1), labels = c("loss", "win"))
```

## Fit the overfitting model

This tree is intentionally allowed to grow aggressively:
- very deep tree
- tiny split restrictions
- no practical pruning

```{r fit-overfit}
overfit_tree <- rpart(
  result ~ .,
  data = train,
  method = "class",
  control = rpart.control(
    maxdepth = 30,
    minsplit = 2,
    minbucket = 1,
    cp = 0
  )
)

printcp(overfit_tree)
```

## Evaluate on training and test sets

```{r evaluate-overfit}
train_pred <- predict(overfit_tree, train, type = "class")
test_pred  <- predict(overfit_tree, test, type = "class")

train_cm <- confusionMatrix(data = train_pred, reference = train$result, positive = "win")
test_cm  <- confusionMatrix(data = test_pred, reference = test$result, positive = "win")

cat("Training confusion matrix:\n")
print(train_cm$table)

cat("\nTesting confusion matrix:\n")
print(test_cm$table)

cat("\nTraining performance summary:\n")
print(train_cm)

cat("\nTesting performance summary:\n")
print(test_cm)
```

## Key values for the paper

```{r paper-values}
cat("Training Accuracy:\n")
print(train_cm$overall["Accuracy"])

cat("\nTraining by-class metrics:\n")
print(train_cm$byClass[c("Sensitivity", "Specificity", "Pos Pred Value", "Neg Pred Value")])

cat("\nTesting Accuracy:\n")
print(test_cm$overall["Accuracy"])

cat("\nTesting by-class metrics:\n")
print(test_cm$byClass[c("Sensitivity", "Specificity", "Pos Pred Value", "Neg Pred Value")])
```

## Notes

This version is intended for extracting the train and test confusion-matrix values and summary metrics for the writeup.
